### 2025-05-06 Chapters/paper1.tex
changes

\begin{figure}[H]
	\centering
	\begin{minipage}[b]{0.45\textwidth}
		\centering
		\includegraphics[width=\textwidth]{img/paper_1/random_forest_confusion_matrix.png}
		\caption{Random Forest Confusion Matrix}
	\end{minipage}
	\hfill
	\begin{minipage}[b]{0.45\textwidth}
		\centering
		\includegraphics[width=\textwidth]{img/paper_1/accuracy_curve.png}
		\caption{Random Forest Accuracy Curve}
	\end{minipage}

	\begin{minipage}[b]{0.45\textwidth}
		\centering
		\includegraphics[width=\textwidth]{img/paper_1/loss_curve.png}
		\caption{Random Forest Loss Curve}
	\end{minipage}
	\hfill
	\begin{minipage}[b]{0.45\textwidth}
		\centering
		\includegraphics[width=\textwidth]{img/paper_1/precision_per_class.png}
		\caption{Random Forest Precision Per Class}
	\end{minipage}
\end{figure}

@@

\begin{figure}[H]
	\centering
	\begin{minipage}[b]{0.45\textwidth}
		\centering
		\includegraphics[width=\textwidth]{img/paper_1/BG_confusion_matrix.png}
		\caption{Bagging Classifier Confusion Matrix}
	\end{minipage}
	\hfill
	\begin{minipage}[b]{0.45\textwidth}
		\centering
		\includegraphics[width=\textwidth]{img/paper_1/accuracy_curve.png}
		\caption{Bagging Classifier Accuracy Curve}
	\end{minipage}

@@

	\begin{minipage}[b]{0.45\textwidth}
		\centering
		\includegraphics[width=\textwidth]{img/paper_1/loss_curve.png}
		\caption{Bagging Classifier Loss Curve}
	\end{minipage}
	\hfill
	\begin{minipage}[b]{0.45\textwidth}
		\centering
		\includegraphics[width=\textwidth]{img/paper_1/precision_per_class.png}
		\caption{Bagging Classifier Precision Per Class}
	\end{minipage}
\end{figure}

@@

\begin{figure}[H]
	\centering
	\begin{minipage}[b]{0.45\textwidth}
		\centering
		\includegraphics[width=\textwidth]{img/paper_1//Ensemble_confusion_matrix.png}
		\caption{Ensemble Learning Confusion Matrix}
	\end{minipage}
	\hfill
	\begin{minipage}[b]{0.45\textwidth}
		\centering
		\includegraphics[width=\textwidth]{img/paper_1//accuracy_curve.png}
		\caption{Ensemble Learning Accuracy Curve}
	\end{minipage}

@@

	\begin{minipage}[b]{0.45\textwidth}
		\centering
		\includegraphics[width=\textwidth]{img/paper_1//loss_curve.png}
		\caption{Ensemble Learning Loss Curve}
	\end{minipage}
	\hfill
	\begin{minipage}[b]{0.45\textwidth}
		\centering
		\includegraphics[width=\textwidth]{img/paper_1//precision_per_class.png}
		\caption{Ensemble Learning Precision Per Class}
	\end{minipage}
\end{figure}

@@

\begin{figure}[H]
	\centering
	\begin{minipage}[b]{0.45\textwidth}
		\centering
		\includegraphics[width=\textwidth]{img/paper_1//confusion_matrix.png}
		\caption{Gradient Boosting Confusion Matrix}
	\end{minipage}
	\hfill
	\begin{minipage}[b]{0.45\textwidth}
		\centering
		\includegraphics[width=\textwidth]{img/paper_1//accuracy_curve.png}
		\caption{Gradient Boosting Accuracy Curve}
	\end{minipage}

@@

	\begin{minipage}[b]{0.45\textwidth}
		\centering
		\includegraphics[width=\textwidth]{img/paper_1//loss_curve.png}
		\caption{Gradient Boosting Loss Curve}
	\end{minipage}
	\hfill
	\begin{minipage}[b]{0.45\textwidth}
		\centering
		\includegraphics[width=\textwidth]{img/paper_1//precision_per_class.png}
		\caption{Gradient Boosting Precision Per Class}
	\end{minipage}
\end{figure}

\section{Results and Evaluation with K-fold Cross Validation }

\begin{figure}[H]
	\centering
	\begin{minipage}[b]{0.45\textwidth}
		\centering
		\includegraphics[width=\textwidth]{img/paper_1/Ensemble Learning_percentage_based.png}
		\caption{Ensemble Learning Confusion Matrix }
	\end{minipage}
	\hfill
	\begin{minipage}[b]{0.45\textwidth}
		\centering
		\includegraphics[width=\textwidth]{img/paper_1/Random Forest_percentage_based.png}
		\caption{Random Forest Confusion Matrix }
	\end{minipage}

	\begin{minipage}[b]{0.45\textwidth}
		\centering
		\includegraphics[width=\textwidth]{img/paper_1/GB Classifier_percentage_based.png}
		\caption{GB Classifier Confusion Matrix }
	\end{minipage}
	\hfill
	\begin{minipage}[b]{0.45\textwidth}
		\centering
		\includegraphics[width=\textwidth]{img/paper_1/SGD_percentage_based.png}
		\caption{SGD Confusion Matrix }
	\end{minipage}
\end{figure}

@@

\begin{figure}[H]
	\centering
	\begin{minipage}[b]{0.45\textwidth}
		\centering
		\includegraphics[width=\textwidth]{img/paper_1/accuracy_comparison.png}
		\caption{Accuracy Comparison}
	\end{minipage}
	\hfill
	\begin{minipage}[b]{0.45\textwidth}
		\centering
		\includegraphics[width=\textwidth]{img/paper_1/bar_chart_improved.png}
		\caption{Improved Bar Chart}
	\end{minipage}

	\begin{minipage}[b]{0.45\textwidth}
		\centering
		\includegraphics[width=\textwidth]{img/paper_1/box_plot_comparison.png}
		\caption{Box Plot Comparison}
	\end{minipage}
	\hfill
	\begin{minipage}[b]{0.45\textwidth}
		\centering
		\includegraphics[width=\textwidth]{img/paper_1/violin_plot_improved.png}
		\caption{Improved Violin Plot}
	\end{minipage}
\end{figure}

>>>>

\begin{figure}[H]
	\centering
	\begin{minipage}[t]{0.48\textwidth}
		\centering
		\includegraphics[width=\textwidth]{img/paper_1/random_forest_confusion_matrix.png}
		\caption{Random Forest Confusion Matrix}
	\end{minipage}
	\hfill
	\begin{minipage}[t]{0.48\textwidth}
		\centering
		\includegraphics[width=\textwidth]{img/paper_1/accuracy_curve.png}
		\caption{Random Forest Accuracy Curve}
	\end{minipage}
\end{figure}
	\vspace{1em}
\begin{figure}[H]
	\begin{minipage}[t]{0.48\textwidth}
		\centering
		\includegraphics[width=\textwidth]{img/paper_1/loss_curve.png}
		\caption{Random Forest Loss Curve}
	\end{minipage}
	\hfill
	\begin{minipage}[t]{0.48\textwidth}
		\centering
		\includegraphics[width=\textwidth]{img/paper_1/precision_per_class.png}
		\caption{Random Forest Precision Per Class}
	\end{minipage}
\end{figure}

The Random Forest classifier demonstrates solid performance with an accuracy of 0.764, precision of 0.777, and sensitivity of 0.770. Its F1-Score of 0.766 reflects a balanced performance between precision and recall.

@@

\begin{figure}[H]
	\centering
	\begin{minipage}[t]{0.48\textwidth}
		\centering
		\includegraphics[width=\textwidth]{img/paper_1/BG_confusion_matrix.png}
		\caption{Bagging Classifier Confusion Matrix}
	\end{minipage}
	\hfill
	\begin{minipage}[t]{0.48\textwidth}
		\centering
		\includegraphics[width=\textwidth]{img/paper_1/accuracy_curve.png}
		\caption{Bagging Classifier Accuracy Curve}
	\end{minipage}

@@

	\begin{minipage}[t]{0.48\textwidth}
		\centering
		\includegraphics[width=\textwidth]{img/paper_1/loss_curve.png}
		\caption{Bagging Classifier Loss Curve}
	\end{minipage}
	\hfill
	\begin{minipage}[t]{0.48\textwidth}
		\centering
		\includegraphics[width=\textwidth]{img/paper_1/precision_per_class.png}
		\caption{Bagging Classifier Precision Per Class}
	\end{minipage}
\end{figure}

The Gradient Boosting classifier achieves moderate performance with an accuracy of 0.702, precision of 0.687, and sensitivity of 0.689. Its F1-Score of 0.687 indicates a fairly balanced trade-off between precision and recall.

@@

\begin{figure}[H]
	\centering
	\begin{minipage}[t]{0.48\textwidth}
		\centering
		\includegraphics[width=\textwidth]{img/paper_1/Ensemble_confusion_matrix.png}
		\caption{Ensemble Learning Confusion Matrix}
	\end{minipage}
	\hfill
	\begin{minipage}[t]{0.48\textwidth}
		\centering
		\includegraphics[width=\textwidth]{img/paper_1/accuracy_curve.png}
		\caption{Ensemble Learning Accuracy Curve}
	\end{minipage}

@@

	\begin{minipage}[t]{0.48\textwidth}
		\centering
		\includegraphics[width=\textwidth]{img/paper_1/loss_curve.png}
		\caption{Ensemble Learning Loss Curve}
	\end{minipage}
	\hfill
	\begin{minipage}[t]{0.48\textwidth}
		\centering
		\includegraphics[width=\textwidth]{img/paper_1/precision_per_class.png}
		\caption{Ensemble Learning Precision Per Class}
	\end{minipage}
\end{figure}

The Ensemble Learning model exhibits strong performance with an accuracy of 0.831, precision of 0.830, and sensitivity of 0.820. Its F1-Score of 0.819 underscores its effectiveness in maintaining a balance between precision and recall.

@@

\begin{figure}[H]
	\centering
	\begin{minipage}[t]{0.48\textwidth}
		\centering
		\includegraphics[width=\textwidth]{img/paper_1/confusion_matrix.png}
		\caption{Gradient Boosting Confusion Matrix}
	\end{minipage}
	\hfill
	\begin{minipage}[t]{0.48\textwidth}
		\centering
		\includegraphics[width=\textwidth]{img/paper_1/accuracy_curve.png}
		\caption{Gradient Boosting Accuracy Curve}
	\end{minipage}

@@

	\begin{minipage}[t]{0.48\textwidth}
		\centering
		\includegraphics[width=\textwidth]{img/paper_1/loss_curve.png}
		\caption{Gradient Boosting Loss Curve}
	\end{minipage}
	\hfill
	\begin{minipage}[t]{0.48\textwidth}
		\centering
		\includegraphics[width=\textwidth]{img/paper_1/precision_per_class.png}
		\caption{Gradient Boosting Precision Per Class}
	\end{minipage}
\end{figure}

The Gradient Boosting Classifier shows competent performance with an accuracy of 0.753, precision of 0.737, and sensitivity of 0.746. Its F1-Score of 0.734 suggests a solid equilibrium between precision and recall.

\section{Results and Evaluation with K-fold Cross Validation}
\begin{table}[!h]
	\centering
	\caption{Classifier Performance Summary}
	\begin{tabular}{lcccc}
		\toprule
		\textbf{Classifier} & \textbf{Accuracy} & \textbf{Precision} & \textbf{Recall} & \textbf{F1-Score} \\
		\midrule
		Ensemble Learning & 0.95 & 0.86 & 0.79 & 0.82 \\
		Random Forest     & 0.88 & 0.85 & 0.81 & 0.82 \\
		Gradient Boosting & 0.87 & 0.83 & 0.81 & 0.82 \\
		SGD Classifier    & 0.80 & 0.73 & 0.70 & 0.70 \\
		\bottomrule
	\end{tabular}
\end{table}

\begin{figure}[H]
	\centering
	\begin{minipage}[t]{0.48\textwidth}
		\centering
		\includegraphics[width=\textwidth]{img/paper_1/Ensemble Learning_percentage_based.png}
		\caption{Ensemble Learning Percentage-Based Confusion Matrix}
	\end{minipage}
	\hfill
	\begin{minipage}[t]{0.48\textwidth}
		\centering
		\includegraphics[width=\textwidth]{img/paper_1/Random Forest_percentage_based.png}
		\caption{Random Forest Percentage-Based Confusion Matrix}
	\end{minipage}
\end{figure}

\textbf{Analysis:} Ensemble Learning outperforms all other models with the highest average accuracy (95\%) and strong macro F1-score (0.82), particularly excelling in the Wake stage with near-perfect precision and recall. Random Forest and Gradient Boosting both deliver balanced performance with 88\% and 87\% accuracy respectively, showing robust results in N2 and REM stages. SGD Classifier, while computationally efficient, lags behind in performance, especially in accurately detecting the N1 stage, reflecting a much lower F1-score (0.70).

\begin{figure}[H]
	\centering
	\begin{minipage}[t]{0.48\textwidth}
		\centering
		\includegraphics[width=\textwidth]{img/paper_1/GB Classifier_percentage_based.png}
		\caption{Gradient Boosting Percentage-Based Confusion Matrix}
	\end{minipage}
	\hfill
	\begin{minipage}[t]{0.48\textwidth}
		\centering
		\includegraphics[width=\textwidth]{img/paper_1/SGD_percentage_based.png}
		\caption{SGD Percentage-Based Confusion Matrix}
	\end{minipage}
\end{figure}

@@

The comparison tables highlight that Gradient Boosting and Ensemble Learning consistently outperform other classifiers across overall and per-stage metrics, particularly in precision and F1-score.

\begin{figure}[H]
	\centering
	\includegraphics[width=0.99\linewidth]{img/paper_1/bar plot}
	\caption{Comparison of Classifier Performance Based on Precision, Recall, and F1-Score}
	\label{fig:bar-plot}
\end{figure}

 Notably, REM and Wake (W) stages achieve the highest precision and recall, while N1 and N2 stages remain more challenging for all models.
\begin{table}[h!]
	\centering
	\begin{tabular}{lcccc}
		\toprule
		\textbf{Classifier} & \textbf{Accuracy} & \textbf{Precision} & \textbf{Sensitivity} & \textbf{F1-Score} \\
		\midrule
		Random Forest       & 0.764             & 0.777              & 0.770                & 0.766             \\
		Gradient Boosting   & 0.702             & 0.687              & 0.689                & 0.687             \\
		Ensemble Learning   & \textbf{0.831}    & \textbf{0.830}     & \textbf{0.820}       & \textbf{0.819}    \\
		Bagging Classifier  & 0.753             & 0.737              & 0.746                & 0.734             \\
		\bottomrule
	\end{tabular}
	\caption{ Performance Metrics of Classifiers}
\end{table}

\begin{table}[htbp]
	\centering

	\begin{tabular}{lcccc}
		\toprule
		\textbf{Stage} & \textbf{Metric} & \textbf{Random Forest} & \textbf{Ensemble (Gradient)} & \textbf{Bagging} \\
		\midrule
		\textbf{W}     & Precision       & 0.86                   & \textbf{0.97}                & 0.95             \\
		& Recall          & \textbf{0.91}          & 0.86                         & \textbf{0.97}    \\
		\textbf{N1}    & Precision       & 0.50                   & \textbf{0.57}                & 0.57             \\
		& Recall          & 0.59                   & \textbf{0.79}                & \textbf{0.79}    \\
		\textbf{N2}    & Precision       & 0.48                   & \textbf{0.68}                & 0.56             \\
		& Recall          & 0.44                   & \textbf{0.56}                & 0.41             \\
		\textbf{REM}   & Precision       & 0.93                   & \textbf{0.98}                & 0.87             \\
		& Recall          & 0.88                   & \textbf{0.97}                & 0.93             \\
		\bottomrule
	\end{tabular}

\end{table}
